\documentclass[oneside]{article} % For LaTeX2e
\usepackage{workshop}
\usepackage{lipsum}
%\usepackage{showframe}
\title{Computing Science Paper Review}
\Instructor{Anna Koop}
\Class{Amii}
\Version{0.5}
\Email{anna@amii.ca}

\begin{document}


\maketitle
%%%%%%%%%%%%%%%%%%% Details
\section*{Paper}\label{sec:ref}
\completeline{Title:}

\completeline{Authors:}

\completeline{Bib info:}

\completeline{Found this reference from:}

%%%%%%%%%%%%%%%%%%% Context
\section*{Project}\label{sec:pro}
\completeline{Goal/Deliverable}

\shortblock{This paper relates to this project by}

\shortblock{The relevant contribution of the paper is}

%\qline,
%%%%%%%%%%%%%%%%%%% Overview

\section*{The 5 Cs}
\label{sec:cs}
\completeline{Category:}

\completeline{Context:}

\completeline{Correctness:}

\completeline{Contribution(s):}

\qline

\qline

\completeline{Clarity:}

%%%%%%%%%%%%%%%%%%% Relevance

\section*{Conclusions}
\label{sec:integration}
\completeline{My overall understanding}

\shortblock{Assumptions}

\shortblock{Questions sparked}

\shortblock{Concepts to follow up on}

\prompt{Relevant references}{3}

\prompt{Actions to take}{3}

\newpage

%%%%%%%%%%%%%%%%%%% Expanded details

\section*{Example Categories}
\label{sec:cat}

\begin{center}
\begin{tabular}{ p{1.5cm} p{6.5cm} p{1.5cm} p{6.5cm}}
 Empirical & \begin{checklist}
 \item evaluation of a novel algorithm
 \item application to a new domain
 \item novel evaluation of existing algorithms
 \item adaptation of existing algorithm
 \end{checklist}&
 Synthesis & \begin{checklist}
 \item comparison of existing methods
 \item case study evaluation
 \item literature survey in subfield
 \item meta-analysis of previous studies
 \end{checklist} \\
Theoretical & \begin{checklist}
\item mathematical proof
\item algorithmic analysis
\item novel application of theory
 \end{checklist} &
Experiential & \begin{checklist}
 \item implementation of a specific system
 \item position paper
 \item tutorial
 \end{checklist}\\
 \end{tabular}
 \end{center}
 
 \section*{Context}
 The source of the paper in terms of industry, field, and authorship. Most important for how it impacts the interpretation of the results/applicability.
 
 \section*{Conclusions}
 Your overall understanding can be rated by which pass of the paper reading you got to:
 \begin{enumerate}
 \item Read title and skimmed abstract.
 \item Read abstract, intro, conclusion. Skimmed section headings, figures, and bibliography.
 \item Read all sections.
 \item Thoroughly read all sections (rare to get this far).
 \end{enumerate}
 
 Note the assumptions made by the authors and how they connect to the assumptions you are maing or would like to make about your project setting.
 

\end{document}
